---
title: "p8105_hw2_zc2443"
output: github_document
author: "Ziyang Chen"
date: 09-27-2019
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

# Problem 1

## Question 1
```{r}
trashwheel = read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel",
                        range = "A2:N338", na = "") %>%
  janitor::clean_names() %>%
  na.omit() %>% 
  mutate(sports_balls = as.integer(round(sports_balls, 0)))
```

## Question 2
```{r}
precip_2017 = read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation",
                        range = "A2:B14", na = "") %>% 
  mutate(Year = 2017)

precip_2018 = read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation",
                        range = "A2:B9", na = "") %>% 
  mutate(Year = 2018)

comb_precip = left_join(precip_2017, precip_2018, by = "Month") %>% 
  mutate(Month = month.name)
```

## Question 3
The trashwheel dataset has `r nrow(trashwheel)` observations and the comb_precip has `r nrow(comb_precip)` observations. For the trashwheel dataset, variable sports ball is the key variable; for the comb_precip dataset, total number of precipitation is the key variable. The total precipitation in 2018 is `r sum(pull(precip_2018,Total))` in. The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trashwheel, year == 2017),sports_balls))`

# Problem 2
```{r}
plos_month = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(col = mon, into = c("year","month","day")) %>%
  select(prez_gop, prez_dem, everything()) %>% 
  pivot_longer(prez_gop:prez_dem, names_to = "president", names_prefix = "prez_") %>% 
  select(-day) %>% 
  mutate(month = recode(month, "01" = "January", "02" = "February", "03" = "March", "04" = "April",
                        "05" = "May", "06" = "June", "07" = "July", "08" = "August", "09" = "September",
                        "10" = "October", "11" = "November", "12" = "December")) %>% 
  filter(value != 0) %>% 
  select(-value)

snp = read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  separate(col = date, into = c("month","day","year")) %>% 
  arrange(as.integer(year), as.integer(month)) %>% 
  select(year, month, close) %>% 
  mutate(month = recode(month, "1" = "January", "2" = "February", "3" = "March", "4" = "April",
                        "5" = "May", "6" = "June", "7" = "July", "8" = "August", "9" = "September",
                        "10" = "October", "11" = "November", "12" = "December"))

unemployment = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "value") %>% 
  mutate(month = recode(month, "Jan" = "January", "Feb" = "February", "Mar" = "March", "Apr" = "April",
                        "May" = "May", "Jun" = "June", "Jul" = "July", "Aug" = "August", "Sep" = "September",
                        "Oct" = "October", "Nov" = "November", "Dec" = "December")) %>%
  mutate(Year = as.character(Year)) %>% 
  select(year = Year, everything())

plos_snp = left_join(plos_month, snp, by = c("year","month"))

plos_snp_unem = left_join(plos_snp, unemployment, by = c("year","month"))
```

The `plos_month` dataset contains the date and the number of each kind of governmental staff. The `snp` dataset contains the date and the closing values of the S&P stock index on the associated date. The `unemploment` dataset contains date(year and month) and the percentage of unemployment corresponding to each month of the associated year. The resulting dataset `plos_snp_unem`has `r nrow(plos_snp_unem)` observations and contains `r ncol(plos_snp_unem)` variables. The year meansure ranges from 1947 to 2015. The key vairables are the `close` and `value`.

# Problem 3
```{r, fig.width=6}
pop_baby = read_csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_upper(gender), ethnicity = str_to_upper(ethnicity),
         childs_first_name = str_to_upper(childs_first_name)) %>% 
  distinct()

olivia = filter(pop_baby, gender == "FEMALE", childs_first_name == "OLIVIA") %>% 
  pivot_wider(names_from = year_of_birth, values_from = rank) %>% 
  select(-gender, -childs_first_name, -count)

pop_male = filter(pop_baby, gender == "MALE", rank == 1) %>% 
  pivot_wider(names_from = year_of_birth, values_from = rank) %>% 
  select(-gender, -ethnicity, -count)

filter(pop_baby, gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>% 
  ggplot(aes(x = rank, y = count)) + geom_point(alpha = 0.4)
```

